Reduced MHD Notes

### 2024-12-29 Notes/Notes.tex
Rename
Edits

For the first term, we have the identity: 
$$(\nabla\times B)\times B = B\cdot\nabla B - \nabla B\cdot B. $$
The components of this expression look like $B_i\p_iB_j - B_j\p_iB_j$ for $i,j\in\{x,y,z\}$. Considering the components separately, we have: 
\begin{equation} \begin{split}
    [(\nabla\times B)\times B]_z &= B\cdot\nabla B_z - (\p_zB)\cdot B \\
    &= (B_\perp\cdot\np B_z + B_z\p_zB_z) - (B_\perp\cdot \p_zB_\perp + B_z\p_zB_z) \\
    &= B_\perp\cdot(\np B_z - \p_zB_\perp) \\ 
    &= \pth{\frac{B_1 \til{B}_0}{a}} \pth{\beta_\perp \cdot \til{\nabla}_\perp\beta_\perp} - \pth{\frac{2\pi B_1^2}{L}} \pth{\p_Z\beta_\perp}. 
\end{split} \end{equation}

@@

\section{Improving Strauss' scaling}
Express the mass density $\rho$, fluid velocity $\bm{v}$, and magnetic field $\bm{B}$ in terms of dimensionless fields $r$, $\bm{\nu}$, $\beta_\parallel$, $\beta_\perp$, according to
\begin{align*}
    \rho &= \rho_0\,(1 + \epsilon\,r)\\
    \bm{v} & = v_0\,\bm{\nu}\\
    \bm{B} & = B_0\bigg((1 + \ep\beta_\para)\bm{e}_z + \ep\bm{\beta}_\perp\bigg) = B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep\beta_\para} .
\end{align*}
Note that the non-constant part of the parallel magnetic field is larger than in Strauss' convention by a single power of $\epsilon$. Also note that the decomposition of density entails assuming small density fluctuations. (That is, to get this system from the previous system, we have only set $\til{B}_0 = B_1 = \ep B_0$ and taken $r\rightarrow (1+\ep r)$. The notation below uses $\np = (\p_X, \p_Y, 0)$).

@@

\subsection{$\nu_\para$ Evolution}
\begin{align} 
    \p_T\nu_\para &= \ep S_0 + \ep S_1 + \ep S_2 \text{, with}
\end{align}
\begin{align*}
    \ep S_0 &= -\ep^2\nu_\para \p_Z\nu_\para - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi}, \\ 
    \ep  S_1 &= -\beta_0\fr\p_Z\pi\pth{\frinv} - \cancel{\ep\p_Z\beta_\para} - \cancel{\ep^2q\p_Z\beta_\para} + \cancel{\ep^2\beta_\para\p_Z\beta_\para} \\ 
        &- \cancel{\ep^2\beta_\para\p_Z\beta_\para} - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 + O\pth{\ep^3}, \\ 
    \ep S_2 &= \cancel{\ep\p_Z\beta_\para} + \cancel{\ep^2 q\p_Z\beta_\para} - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} \\ 
        &- \ep\br{\beta_\para, \Psi} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi} + O\pth{\ep^3}. 
\end{align*}
Only one term changes in different regimes, namely 
\begin{align*}
    -\beta_0\fr\p_Z\pi\pth{\frinv} &= - \beta_0\ep\pi'(1) \p_Z\beta_\para + \beta_0\ep\pi'(1) \p_Zq \\ 
        &- \beta_0\ep^2\pi'(1) \pth{\cancel{q}-\beta_\para} \p_Z\beta_\para + \cancel{\beta_0\ep^2\pi'(1) \pth{q-\beta_\para} \p_Zq} \\ 
        &- \beta_0\ep^2\pi''(1) \beta_\para\p_Z \beta_\para + \beta_0\ep^2\pi''(1) \beta_\para\p_Zq + \beta_0\ep^2\pi''(1) q\p_Z\beta_\para - \beta_0\ep^2\pi''(1) q\p_Z q \\ 
        &+ \cancel{\beta_0\ep^2\pi'(1) \beta_\para\p_Zq} - \cancel{2}\beta_0\ep^2\pi'(1) q\p_Z q + \cancel{\beta_0\ep^2\pi'(1) q\p_Z\beta_\para} + O\pth{\ep^3}. 
\end{align*}
There are no explicit $O(1)$ contributions. Considering cancellations between different $S$ terms, the only $O(\ep)$ contributions are
\begin{align*}
    \ep S_0 &: - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi} \\ 
    \ep S_2 &: - \ep\br{\beta_\para, \Psi}. 
\end{align*}
At $O\pth{\ep^2}$, we have 
\begin{align*}
    \ep S_0 &: -\ep^2\nu_\para \p_Z\nu_\para \\ 
    \ep S_1 &: - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 \\ 
    \ep S_2 &: - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi}
\end{align*}

@@

\begin{equation}
    Df(x): U\to Hom(U,E)
\end{equation}

>>>>

For the first term, we have the identity: 
$$(\nabla\times B)\times B = B\cdot\nabla B - \nabla B\cdot B. $$
The components of this expression look like $B_i\p_iB_j - B_j\p_iB_j$ for $i,j\in\{x,y,z\}$. Considering the components separately, we have: 
\begin{equation} \begin{split}
    [(\nabla\times B)\times B]_z &= B\cdot\nabla B_z - (\p_zB)\cdot B \\
    &= (B_\perp\cdot\np B_z + B_z\p_zB_z) - (B_\perp\cdot \p_zB_\perp + B_z\p_zB_z) \\
    &= B_\perp\cdot(\np B_z - \p_zB_\perp) \\ 
    &= \pth{\frac{B_1 \til{B}_0}{a}} \pth{\beta_\perp \cdot \til{\nabla}_\perp\beta_\perp} - \pth{\frac{2\pi B_1^2}{L}} \pth{\p_Z\beta_\perp} (wrong). 
\end{split} \end{equation}

@@

\section{Improving Strauss' scaling}
Express the mass density $\rho$, fluid velocity $\bm{v}$, and magnetic field $\bm{B}$ in terms of dimensionless fields $r$, $\bm{\nu}$, $\beta_\parallel$, $\beta_\perp$, according to
\begin{align*}
    \rho &= \rho_0\,(1 + \epsilon\,r)\\
    \bm{v} & = v_0\,\bm{\nu}\\
    \bm{B} & = B_0\bigg((1 + \ep\beta_\para)\bm{e}_z + \ep\bm{\beta}_\perp\bigg) = B_0 \pmat{\ep\beta_x \\ \ep\beta_y \\ 1+\ep\beta_\para} .
\end{align*}
Note that the non-constant part of the parallel magnetic field is larger than in Strauss' convention by a single power of $\epsilon$. Also note that the decomposition of density entails assuming small density fluctuations. (That is, to get this system from the previous system, we only have to set $\til{B}_0 = B_1 = \ep B_0$ and take $r\rightarrow (1+\ep r)$. The notation below uses $\np = (\p_X, \p_Y, 0)$).

@@

\subsection{$\nu_\para$ Evolution}
\begin{align} 
    \p_T\nu_\para &= \ep S_0 + \ep S_1 + \ep S_2 \text{, with}
\end{align}
\begin{align*}
    \ep S_0 &= -\ep^2\nu_\para \p_Z\nu_\para - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi}, \\ 
    \ep  S_1 &= -\beta_0\fr\p_Z\pi\pth{\frinv} - \cancel{\ep\p_Z\beta_\para} - \cancel{\ep^2q\p_Z\beta_\para} + \cancel{\ep^2\beta_\para\p_Z\beta_\para} \\ 
        &- \cancel{\ep^2\beta_\para\p_Z\beta_\para} - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 + O\pth{\ep^3}, \\ 
    \ep S_2 &= \cancel{\ep\p_Z\beta_\para} + \cancel{\ep^2 q\p_Z\beta_\para} - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} \\ 
        &- \ep\br{\beta_\para, \Psi} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi} + O\pth{\ep^3}. 
\end{align*}
Only one term changes in different regimes, namely 
\begin{align*}
    -\beta_0\fr\p_Z\pi\pth{\frinv} &= - \beta_0\ep\pi'(1) \p_Z\beta_\para + \beta_0\ep\pi'(1) \p_Zq \\ 
        &- \beta_0\ep^2\pi'(1) \pth{\cancel{q}-\beta_\para} \p_Z\beta_\para + \cancel{\beta_0\ep^2\pi'(1) \pth{q-\beta_\para} \p_Zq} \\ 
        &- \beta_0\ep^2\pi''(1) \beta_\para\p_Z \beta_\para + \beta_0\ep^2\pi''(1) \beta_\para\p_Zq + \beta_0\ep^2\pi''(1) q\p_Z\beta_\para - \beta_0\ep^2\pi''(1) q\p_Z q \\ 
        &+ \cancel{\beta_0\ep^2\pi'(1) \beta_\para\p_Zq} - \cancel{2}\beta_0\ep^2\pi'(1) q\p_Z q + \cancel{\beta_0\ep^2\pi'(1) q\p_Z\beta_\para} + O\pth{\ep^3}. 
\end{align*}
There are no explicit $O(1)$ contributions. Considering cancellations between different $S$ terms, the only $O(\ep)$ contributions are
\begin{align*}
    \ep S_0 &: - \ep\Braket{\nu_\para, \phi} + \ep\br{\nu_\para, \psi} \\ 
    \ep S_2 &: - \ep\br{\beta_\para, \Psi}. 
\end{align*}
At $O\pth{\ep^2}$, we have 
\begin{align*}
    \ep S_0 &: -\ep^2\nu_\para \p_Z\nu_\para \\ 
    \ep S_1 &: - \frac{1}{2}\ep^2\p_Z |\np\psi|^2 = -\ep^2\np\psi \cdot \p_Z\np\psi = -\ep^2 \Braket{\psi, \p_Z\psi} \\ 
    \ep S_2 &: - \ep^2\Braket{\beta_\para, \GN\p_Z\beta_\para} - \ep^2q\br{\beta_\para, \Psi} + \ep^2\beta_\para \br{\beta_\para, \Psi}
\end{align*}

@@

\begin{equation}
    Df(x): U\to\text{Hom}(U,E)
\end{equation}
